---
title: "Lab 5: Student Evaluations of Teaching"
author: "Your name here!"
format: gfm
editor: source
execute: 
  echo: true
---

In this lab, we will be using the `dplyr` package to explore student evaluations
of teaching data. 

**You are expected to use functions from `dplyr` to do your data manipulation!**

## The Data

The `teacher_evals` dataset contains student evaluations of reaching (SET)
collected from students at a University in Poland. There are SET surveys from 
students in all fields and all levels of study offered by the university.

The SET questionnaire that every student at this university completes is as
follows:

> Evaluation survey of the teaching staff of University of Poland. Please
> complete the following evaluation form, which aims to assess the lecturer’s
> performance. Only one answer should be indicated for each question. The
> answers are coded in the following way: 5 - I strongly agree; 4 - I agree;
> 3 - Neutral; 2 - I don’t agree; 1 - I strongly don’t agree.
>
> Question 1: I learned a lot during the course.
>
> Question 2: I think that the knowledge acquired during the course is very
> useful.
>
> Question 3: The professor used activities to make the class more engaging.
>
> Question 4: If it was possible, I would enroll for a course conducted by this
> lecturer again.
>
> Question 5: The classes started on time.
>
> Question 6: The lecturer always used time efficiently.
>
> Question 7: The lecturer delivered the class content in an understandable and
> efficient way.
>
> Question 8: The lecturer was available when we had doubts.
>
> Question 9. The lecturer treated all students equally regardless of their
> race, background and ethnicity.

These data are from the end of the winter semester of the 2020-2021 academic
year. In the period of data collection, all university classes were entirely
online amid the COVID-19 pandemic. While expected learning outcomes were not
changed, the online mode of study could have affected grading policies and could
have implications for data.

**Average SET scores** were combined with many other variables, including:

1.  **characteristics of the teacher** (degree, seniority, gender, SET scores in
the past 6 semesters).
2.  **characteristics of the course** (time of day, day of the week, course
type, course breadth, class duration, class size).
3.  **percentage of students providing SET feedback.**
4.  **course grades** (mean, standard deviation, percentage failed for the
current course and previous 6 semesters).

This rich dataset allows us to **investigate many of the biases in student evaluations of teaching** that have been reported in the literature and to formulate new
hypotheses.

Before tackling the problems below, study the description of each variable
included in the `teacher_evals_codebook.pdf`.

**1. Load the appropriate R packages for your analysis.**

```{r}
#| label: setup
# code chunk for loading packages

```

**2. Load in the `teacher_evals` data.** 

```{r}
#| label: load-data
# code chunk for importing the data

```

### Data Inspection + Summary

**3. Provide a brief overview (~4 sentences) of the dataset.**

```{r}
#| label: explore-data
# you may want to use code to answer this question

```

**4. What is the unit of observation (i.e. a single row in the dataset) identified by?**

```{r}
#| label: row-identification
# you may want to use code to answer this question

```

**5. Use _one_ `dplyr` pipeline to clean the data by:**

- **renaming the `gender` variable `sex`**
- **removing all courses with fewer than 10 respondents**
- **changing data types in whichever way you see fit (e.g., is the instructor ID really a numeric data type?)**
- **only keeping the columns we will use -- `course_id`, `teacher_id`, `question_no`, `no_participants`, `resp_share`, `SET_score_avg`, `percent_failed_cur`, `academic_degree`, `seniority`, and `sex`**

**Assign your cleaned data to a new variable named `teacher_evals_clean` –- use these data going forward. Save the data as `teacher_evals_clean.csv` in the `data-clean` folder.**

```{r}
#| label: data-cleaning
# code chunk for Q4

```

**6. How many unique instructors and unique courses are present in the cleaned dataset?**

```{r}
#| label: unique-courses
# code chunk for Q5

```

**7. One teacher-course combination has some missing values, coded as `NA`. Which instructor has these missing values? Which course? What variable are the missing values in?**

```{r}
#| label: uncovering-missing-values
# code chunk for Q6

```

**8. What are the demographics of the instructors in this study? Investigate the variables `academic_degree`, `seniority`, and `sex` and summarize your findings in ~3 complete sentences.**

```{r}
#| label: exploring-demographics-of-instructors
# code chunk for Q7

```

**9. Each course seems to have used a different subset of the nine evaluation questions. How many teacher-course combinations asked all nine questions?**

```{r}
#| label: teacher-course-asked-every-question
# code chunk for Q8

```

## Rate my Professor

**10. Which instructors had the highest and lowest average rating for Question 1 (I learned a lot during the course.) across all their courses?**

```{r}
#| label: question-1-high-low
# code chunk for Q9

```

**11. Which instructors with one year of experience had the highest and lowest average percentage of students failing in the current semester across all their courses?**

```{r}
#| label: one-year-experience-failing-students
# code chunk for Q10

```

**12. Which female instructors with either a doctorate or professional degree had the highest and lowest average percent of students responding to the evaluation across all their courses?**

```{r}
#| label: female-instructor-student-response
# code chunk for Q11

```
