---
title: "PA 5.2: Jewel Heist"
author: "The names of your group members go here!"
format: html
embed-resources: true
code-tools: true
toc: true
editor: source
execute: 
  error: true
  echo: true
  message: false
  warning: false
---

```{r}
#| label: setup

library(tidyverse)
```

## Solve the Mystery

Just down the road in Montecito, CA several rare jewels went missing. The jewels
were stolen and replaced with fakes, but detectives have not been able to solve
the case. They are now calling in a data scientist to help parse their clues.

A camera was located near the building where the jewels went missing, so the
detectives have provided you with a list of people who may have entered the
building. This list includes the date and time they were spotted on the camera,
**in Pacific Standard Time (PST)**.

Unfortunately, the date and time of the jewel heist is not known. You have been
hired to crack the case. Use the clues below to discover the thief's identity.

```{r}
#| label: read-in-data

suspects <- read_csv("https://raw.githubusercontent.com/zoerehnberg/STAT331-S23/main/practice_activities/suspects.csv")

dim(suspects)
# 214 total suspects

```

<!-- The person who has broken the most bones starts as the Typer (listening to explanations from the Talker)!  -->

1. The crime was committed in Montecito, CA, so we need to make sure the time
zone for these dates are correct. Modify the `Time.Spotted` variable to have the
correct time zone. 

```{r}
#| label: set-correct-tz

```

2. Based on the cleaning schedule for the room where the jewels are held, the
heist was not committed in the morning.

```{r}
#| label: crime-not-committed-in-morning
# After this step you should have 112 suspects left


```

<!-- Swap roles -- Developer becomes Coder, Coder becomes Developer! -->

3. The room where the heist was committed is closed on Tuesdays and Thursdays
(and there were no signs of forced entry), so the heist did not happen on those
days.

```{r}
#| label: crime-not-committed-on-tuesday-or-thursday
# After this step you should have 78 suspects left


```

4. It is believed that the heist was committed within 5 weeks (35 days) of
Thanksgiving 2022. *Hint: Pay attention to time zones!*

```{r}
#| label: crime-committed-within-5-days-of-thanksgiving
# After this step you should have 11 suspects left


```

<!-- Swap roles -- Developer becomes Coder, Coder becomes Developer! -->

5. The detectives partially decoded a message from the thief to a notorious
fence in Iceland. In it, the thief said the job would be done "after the sun
sets for you, but before midnight." 
*Hint: When does the sun set in Iceland at that time of year?*

```{r}
#| label: crime-committed-after-sunset-before-midnight
# After this step you should have 4 suspects left


```

6. The thief left behind a receipt at the scene of the crime. The receipt is
smudged, but the day of the month is shown to be 22. It is thought that the
heist took place no more than three days after the receipt was issued.

```{r}
#| label: crime-committed-within-three-days-of-the-22nd
# After this step you should have 2 suspects left


```

<!-- Swap roles -- Developer becomes Coder, Coder becomes Developer! -->

7. The thief is amused by your efforts and has sent you a cryptic clue:

> "The exact number of seconds between midnight on Jan 1, 1970 and the time
> I arrived on the scene is divisible by 6."

```{r}
#| label: crime-committed-message
# You should end with 1 suspect left


```

## Canvas Quiz Submission

Who is the thief? *Only one name should remain.*
