---
title: "PA 8: United Nations Voting Records"
author: "Instructions"
format: html
code-tools: true
toc: true
editor: source
execute: 
  echo: true
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(lubridate) #should already be installed as a part of the tidyverse, we will talk more about this in Week 10
library(scales)  #you may need to install package `scales`
library(gt) #for tables
library(janitor)
```


***This task is complex. It requires many different types of abilities. Everyone will be good at some of these abilities but nobody will be good at all of them. In order to solve this puzzle, you will need to use the skills of each member of your group.***

# Groupwork Protocols

During the Practice Activity, you and your partner will alternate between two roles---Computer and Coder.

When you are the **Computer**, you will type into the Quarto document in RStudio. However, you **do not** type your own ideas. Instead, you type what the Coder tells you to type. You are permitted to ask the Coder clarifying questions, and, if both of you have a question, you are permitted to ask the professor. You are expected to run the code provided by the Coder and, if necessary, to work with the Coder to debug the code. Once the code runs, you are expected to collaborate with the Coder to write code comments that describe the actions taken by your code.

When you are the **Coder**, you are responsible for reading the instructions / prompts and directing the Computer what to type in the Quarto document. You are responsible for managing the resources your group has available to you (e.g., cheatsheet, textbook). If necessary, you should work with the Computer to debug the code you specified. Once the code runs, you are expected to collaborate with the Computer to write code comments that describe the actions taken by your code.

Here are more details of the [Pair Programming Protocols](../pair-programming-norms.qmd)

::: {.callout-note}
The partner was born the furthest away from CSUMB will start as the Computer (typing and listening to instructions from the Coder). 
:::

### Group Norms

Remember, your group is expected to adhere to the following norms:  

1. Be curious. Don't correct. 
2. Be open minded.
3. Ask questions rather than contribute. 
4. Respect each other.
5. Allow each teammate to contribute to the activity through their role. 
6. Do not divide the work.
7. No cross talk with other groups. 
8. Communicate with each other!

## Goals for the Activity  

- Join multiple data tables together by a common variable(s)  
- Create new data sets through the joining of data from various sources  
- Combine `join` functions with other `tidyverse` functions  

**THROUGHOUT THE Activity** be sure to follow the Style Guide by doing the following:  

- load the appropriate packages at the beginning of the Quarto document  
- use proper spacing  
- *add labels* to all code chunks  
- comment at least once in each code chunk to describe why you made your coding decisions  
- add appropriate labels to all graphic axes  


## Setting up your Project

Your project should have the following components:

1. completed [pa-8-united-nations-voting-activity.qmd](../week-9/pa-8-united-nations-voting-activity.qmd)
2. rendered file as `.html`

::: {.callout-important}
The original `Computer` should submit the zip file for the canvas quiz. The original `Coder` can just submit the rendered html file. 

`Computer` - Be sure to share the final .qmd and .html file with the original `Coder`.
:::




# Data Description
The data this week comes from Harvard's Dataverse by way of Mine Çetinkaya-Rundel, David Robinson, and Nicholas Goguen-Compagnoni.

> Original Data citation: Erik Voeten "Data and Analyses of Voting in the UN General Assembly" Routledge Handbook of International Organization, edited by Bob Reinalda (published May 27, 2013). Available at SSRN: http://ssrn.com/abstract=2111149

It was featured on [TidyTuesday](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-03-23/readme.md)

Here is each data set and its description (you might want to look at the Tidy Tuesday link for the tables already rendered)

`unvotes.csv`

|variable     |class     |description |
|:------------|:---------|:-----------|
|rcid         |double    | The roll call id |
|country      |character | Country name, by official English short name |
|country_code |character | 2-character ISO country code |
|vote         |integer   | Vote result as a factor of yes/abstain/no |


```{r}
#| label: read-unvotes
#| echo: false
unvotes <- read_csv("data/unvotes.csv", show_col_types = FALSE)
slice_sample(unvotes, n = 10) |> 
  gt() |> 
  tab_header(title = "unvotes")
```



`roll_calls.csv`

|variable      |class     |description |
|:-------------|:---------|:-----------|
|rcid          |integer   |The roll call id |
|session       |double    | Session number. The UN holds one session per year; these started in 1946|
|importantvote |integer   | Whether the vote was classified as important by the U.S. State Department report "Voting Practices in the United Nations". These classifications began with session 39|
|date          |double    | Date of the vote, as a Date vector|
|unres         |character | Resolution code |
|amend         |integer   | Whether the vote was on an amendment; coded only until 1985 |
|para          |integer   | Whether the vote was only on a paragraph and not a resolution; coded only until 1985|
|short         |character |  Short description |
|descr         |character | Longer description|


```{r}
#| label: read-roll_calls
#| echo: false
#edit to read in roll_calls
roll_calls <- read_csv("data/roll_calls.csv", show_col_types = FALSE)
slice_sample(roll_calls, n = 10) |> 
  gt() |> 
  tab_header(title = "roll_calls")
```



`issues.csv`

|variable   |class     |description |
|:----------|:---------|:-----------|
|rcid       |integer   | The roll call id|
|short_name |character | Two-letter issue codes |
|issue      |integer   | Descriptive issue name |

```{r}
#| label: read-issues
#| echo: false
#edit to read in issues
issues <- read_csv("data/issues.csv", show_col_types = FALSE)
slice_sample(issues, n = 10) |>   
  gt() |> 
  tab_header(title = "issues")
```


**What variable(s) do each of these data frames have in common?**




<!-- Swap roles -- Computer becomes Coder, Coder becomes Computer! -->



## Data Exploration
Our goal today is to explore how various members (countries) in the United Nations vote. We have three data sets, what can we determine from each data set separately?


### UN Votes

The first data set, `unvotes` contains data on the `rcid` which is the roll call id for the vote, the country/country code, and how the country voted. What can we learn from the data?  


Comment on the following code, what is happening in each line?  One way to approach seeing what each line does is to highlight the code from before the `pipe` of that line up to the data `unvotes` and use `CTRL + ENTER` to run just the highlighted lines.  

```{r}
#| label: unvotes-graph
unvotes |> 
  count(country, vote) |> #comment
  group_by(country) |> #comment
  mutate(total = sum(n)) |> #comment 
  mutate(prop_vote = n/total) |> #comment
  filter(country %in% c("United States", "Canada", 
                        "Germany", "France",
                        "Italy", "Japan",
                        "United Kingdom")) |> #comment
  ggplot(aes(y = country, x = prop_vote, 
             fill = vote)) + #comment
  geom_col(position = position_stack()) + #comment
  labs(y = "Group of Seven Countries",
       x = "Proportion of Votes",
       title = "Voting Record of the G7",
       fill = "Vote") + #comment
  theme_minimal() +  #comment
  scale_fill_viridis_d(end = 0.8) 
```

**Describe what the graph above demonstrates above UN voting records for the G7.** 
  
  
  
  

<!-- Swap roles -- Computer becomes Coder, Coder becomes Computer! -->




### Roll Calls

The second data set, `roll_calls` has more information on the type of vote, the importance, whether it was a resolution, and date of the vote.  

```{r}
#| label: roll-call-list
roll_calls |> 
  distinct(short)
```


**What does the code above do? What information does it provide? Is it useful?**
> Description of code results



We can use the individual data for `roll_calls` to look at the number of votes per year over time.

```{r}
#| label: roll-call-graph
roll_calls |> 
  mutate(year = lubridate::year(date)) |> #extracts the year from the date value and creates a new `year` column
  count(year) |> #counts how many votes there were per year assuming each line is an single voting instance
  ggplot(aes(x = year, y = n)) +
  geom_line() +
  labs(x = "Year", y = "Number of Votes",
       title = "UN Votes per Year") +
  theme_minimal()
```



**What information is missing from the above graphic that might be useful in understanding the issues the UN commonly votes on?**





<!-- Swap roles -- Computer becomes Coder, Coder becomes Computer! -->


### Issues 

Finally we have the `issues` data which provides a more general description for each vote on specific issues.  Note that not all issues are included in the data set, just the ones related to the 6 issues below:

```{r}
#| label: count-issues
issues |> 
  count(issue) |> 
  adorn_totals("row") #from janitor package
```

Notice the size of the issues data - it has 5745 rows, but if look at the distinct number of roll call identification numbers, you will see that there are 4099, meaning that more than one issue can be associated with the same roll call id/vote. 

```{r}
#| label: distinct-issues-rcid
issues |> 
  distinct(rcid) |> 
  n_distinct()
```

Recall that in `roll_call` there are 6202 distinct roll call ids/votes, so the issues associated with `issues` do not represent all votes (i.e., there were other U.N. votes on other issues than our 6 chosen issues).  

```{r}
#| label: distinct-issues-rcid-n
roll_calls |> 
  distinct(rcid) |> 
  n_distinct()
```


It would be helpful to use the `issues` data with the `roll_calls` data to be able to better understand the voting trends within the UN on these 6 issues. To do this, we need to join the data.  



# Votes Over Time  

Now let's join our data together to get a better idea of how the UN has voted over time.  
First, look at the number of rows in `issues` and `roll_calls` - do they match? What does this indicate?

```{r}
#| label: check-dimensions
dim(issues)
dim(roll_calls)
```


Now let's try joining the `roll_calls` with the `issues` data.  Compare the following codes to join the data. Describe what each one does and how it differs from the others as a comment in the code chunk.  You might need to reference the slides or reading for this week.

```{r}
#| label: left-join
#| message: true
#| warning: true
#| eval: false
roll_calls |> 
  left_join(issues, by = "rcid") #description of join
```

```{r}
#| label: right-join
#| message: true
#| warning: true
#| eval: false
roll_calls |> 
  right_join(issues, by = "rcid") #description of join
```

```{r}
#| label: full-join
#| message: true
#| warning: true
#| eval: false
roll_calls |> 
  full_join(issues, by = "rcid") #description of join
```

```{r}
#| label: inner-join
#| message: true
#| warning: true
#| eval: false
roll_calls |> 
  inner_join(issues, by = "rcid") #description of join
```


**If we are only interested in retaining the records associated with the `issues` labeled in that data frame, ignoring the other votes, which join should we use?**



<!-- Swap roles -- Computer becomes Coder, Coder becomes Computer! -->


Now that we know how to join the data, we will use the following code to examine the the voting trends for three of the issues related to conflict/weapons. 

Be sure to run the code via the green arrow on the code chunk, as the `case_when()` code can get finicky sometimes and claim an error about a comma in the code when it doesn't exist.  *Comment on the code where indicated* and add your chosen join function


```{r}
#| label: votes-on-conflicts-graph
#| eval: false
roll_calls |> 
 __________join(issues, by = "rcid") |>  #join roll_calls and issues so that just the votes related to the issues data are retained.
  mutate(issue_short = case_when(
    issue == "Arms control and disarmament" ~ "Arms Control",
    issue == "Nuclear weapons and nuclear material" ~ "Nuclear Weapons",
    issue == "Palestinian conflict" ~ "Palestinian Conflict",
    TRUE ~ issue)) |>   #what is happening in this mutate function?
  filter(issue_short %in% c("Arms Control",
                            "Nuclear Weapons",
                            "Palestinian Conflict")) |> #what are we doing here?
  mutate(year = lubridate::year(date)) |> #create a column `year` that contains the year value
  count(year, 
        issue_short) |> #what does this line do?
  ggplot(aes(x = year, 
             y = n, 
             group = issue_short)) + #what does this line do?
  geom_line(aes(color = issue_short)) + #what does this line do?
  labs(x = "Year", 
       y = "Number of Votes",
       title = "United Nations Votes per Year",
       subtitle = "Conflict and Arms Related Votes",
       color = "Voting Issue") +
  theme_minimal() +
  scale_color_viridis_d(end = 0.9)
```


```{r}
#| label: votes-on-conflicts-graph-correct
#| echo: false
roll_calls |> 
 right_join(issues, by = "rcid") |>  #join roll_calls and issues so that just the votes related to the issues data are retained.
  mutate(issue_short = case_when(
    issue == "Arms control and disarmament" ~ "Arms Control",
    issue == "Nuclear weapons and nuclear material" ~ "Nuclear Weapons",
    issue == "Palestinian conflict" ~ "Palestinian Conflict",
    TRUE ~ issue)) |>   #simplifies the issue code for three issues
  filter(issue_short %in% c("Arms Control",
                            "Nuclear Weapons",
                            "Palestinian Conflict")) |> #filters the 3 codes of interest
  mutate(year = lubridate::year(date)) |> #create a column `year` that contains the year value
  count(year, 
        issue_short) |> #counts up the issue votes by year?
  ggplot(aes(x = year, 
             y = n, 
             group = issue_short)) + #connects the lines by issue code
  geom_line(aes(color = issue_short)) + #colors lines by issue code
  labs(x = "Year", 
       y = "Number of Votes",
       title = "United Nations Votes per Year",
       subtitle = "Conflict and Arms Related Votes",
       color = "Voting Issue") +
  theme_minimal() +
  scale_color_viridis_d(end = 0.9)
```


**What do you notice? What do you wonder based on the graph created?**




<!-- Swap roles -- Computer becomes Coder, Coder becomes Computer! -->


# Joining all Data

We want to try create a visualization that compares the voting records of the US and Canada on the 6 major issues of interest. To do this, though, we need information from all three data set, `unvotes`, `issues`, and `roll_calls`

We want to join all three data sets together, *maintaining only the votes for which we have identified the general issue* (e.g., Nuclear War, Arms, Economics, etc.), but recognizing that each `rcid` will match MULTIPLE rows in the `unvotes` because we have each individual country's vote.  We will save (assign) the data as `un_full`.  

```{r}
#| label: join-all
un_full <- roll_calls |> 
  right_join(issues, by = "rcid") |> 
  left_join(unvotes, by = "rcid", 
            multiple = "all", 
            relationship = "many-to-many") 
```


**Describe what each join is doing and why each join has specific arguments.**



::: {.callout-note}
It will be helpful to look up the arguments in the `left_join()` function on the [`dplyr` webpage](https://dplyr.tidyverse.org/reference/mutate-joins.html). 
:::


Now, we are going to do some data cleaning.  Our goal is create a data set that includes the percentage of "yes" votes per country each year. We will call the data table `yes_votes`. *Provide a comment to describe what each line of code is doing in the process*.  

```{r}
#| label: yes-votes-cleaning
yes_votes <- un_full |> 
  select(country, 
         issue, 
         date, 
         vote) |>  # your comment here
  mutate(year = lubridate::year(date)) |> #create a new variable called year
  group_by(country, 
           year, 
           issue) |> # your comment here
  summarize(prop_yes = mean(vote == "yes"), 
            .groups = "drop_last") |> #calculate the proportion of yes votes
  mutate(issue = case_when(
    issue == "Arms control and disarmament" ~ "Arms Control",
    issue == "Nuclear weapons and nuclear material" ~ "Nuclear Weapons",
    issue == "Palestinian conflict" ~ "Palestinian Conflict",
    TRUE ~ issue)) # your comment here
```


```{r}
#| label: yes-votes-cleaning-correct
#| echo: false
yes_votes <- un_full |> 
  select(country, 
         issue, 
         date, 
         vote) |>  #keep just the columns for country, issue, date, vote
  mutate(year = lubridate::year(date)) |> #create a new variable called year
  group_by(country, 
           year, 
           issue) |> #group country, year, issue together to prepare for subsequent analysis
  summarize(prop_yes = mean(vote == "yes"), 
            .groups = "drop_last") |> #calculate the proportion of yes votes
  mutate(issue = case_when(
    issue == "Arms control and disarmament" ~ "Arms Control",
    issue == "Nuclear weapons and nuclear material" ~ "Nuclear Weapons",
    issue == "Palestinian conflict" ~ "Palestinian Conflict",
    TRUE ~ issue)) #create/overwrite issue column with simpler values

#View(yes_votes) #you can use this code to look at the table created
```


Now we can feed the `yes_votes` transformed data table into your graphing code, but first we will want to focus on the United States and Canada. *Provide a comment to describe what each line of code is doing in the process*.  

```{r}
#| label: yes-votes-canada-us
#| message: false
yes_votes |> 
  filter(country %in% c("United States","Canada")) |> # your comment here
  ggplot(mapping = aes(x = year, y = prop_yes, color = country)) + # your comment here
  geom_point(alpha = 0.4) + # your comment here
  #geom_line(aes(group = country)) +
  geom_smooth(method = "loess", se = FALSE) + #this fits a special model called a loess regression, a smooth line that fits the data
  facet_wrap(~issue) + # your comment here
  scale_y_continuous(labels = scales::percent) + #your comment here
  labs(
    title = "Percentage of 'Yes' votes in the UN General Assembly",
    subtitle = "1946 to 2019",
    y = "% Yes",
    x = "Year",
    color = "Country"
  ) + #your comment here
  theme_bw() +
  theme(legend.position = "bottom") + #your comment here 
  scale_color_viridis_d(option = "turbo") #your comment here
```


**What do you notice about the voting records over time?**


# Demonstration: Adding more data
After your instructor created the above plot, she became curious about how politics might impact the UN Voting record for the United States since UN Ambassador is a presidential appointment. So your instructor started searching for a data set of US presidents, their years in office, and their political affiliation.  She found a data set on Kaggle.com and removed the information prior to 1940 (because the dates were coded funny and it was causing problems). She saved the data as `us_presidents.csv` and imported it into the project.

```{r}
#| label: read-president
president <- read_csv("data/us_presidents.csv", show_col_types = FALSE)
slice_sample(president, n = 10) |>   
  gt() |> 
  tab_header(title = "president")
```

She realized that her data only had the start/end dates for each president and she wanted a data set that filled in the missing years and political party for the president in that time period.  After much googling and reading Stack Overflow, found two functions she did not know about called `complete()` and `fill()` to fill in the missing years and party affiliations  


```{r}
#| label: filling-in-year

politics_year <- president |> 
    mutate(start = lubridate::mdy(start),  #formats date correctly
           start_year = lubridate::year(start)) |> #pulls out year
  filter(start_year > 1940) |>  #removes data before 1940 since there was no UN
  select(start_year, party) |>   #pulls out just the variables of interest
  complete(start_year = seq(min(start_year), 2020, by = 1)) |> #fill in missing years
  fill(party) #fill in missing party affiliations for years

slice_tail(politics_year, n = 10) |>   
  gt() |> 
  tab_header(title = "Year by Presidential Party")

```


Next, your instructor took the `yes_votes` data and filtered out just the United States data and then joined by year to add the party affiliation of the president for each year of UN votes.  To create the visualization with the smoothed model, but color by party affiliation, she had to add a new column called `predict` that fit the model first instead of using `geom_smooth()`.  

```{r}
#| label: full-clean-president-graph
yes_votes |> 
  filter(country == "United States") |>  #we want to focus on the US
  left_join(politics_year, 
            by = c("year" = "start_year")) |> #adding in president political affiliation data
  group_by(issue) |>  #we want to calculate predicted yes by issue type 
  mutate(predict = predict(loess(prop_yes ~ year))) |>  #creates predictions using a loess model for yes vs year
  ungroup() |> #ungroups the group_by line so it doesn't mess with other calculations
  ggplot(mapping = aes(x = year, 
                       color = party, 
                       group = 1)) + #sets universal aes to observe changes over time
  geom_point(aes(y = prop_yes),
             alpha = 0.4) + #plots points for proportion yes but transparent
  geom_line(aes(y = predict)) + #plots the predicted yeses for votes
  facet_wrap(~issue) + #creates a plot for each issue
  scale_y_continuous(labels = scales::percent) + #scales y-axis as percent values
  labs(
    title = "Percentage of United States 'Yes' votes in the UN General Assembly",
    subtitle = "1946 to 2019",
    y = "% Yes",
    x = "Year",
    color = "Presidential Political Party"
  ) + #adds labels
  theme_bw() + #changes appearance
  scale_color_manual(values = c("blue", "red"),
                     breaks = c("Democratic", "Republican")) + #sets specific colors for each party
  theme(legend.position = "bottom")  #moves legend to bottom to create space
```
