---
title: "PA 3: Using Data Visualization to Find the Penguins"
format: html
editor: source
embed-resources: true
execute: 
  echo: true
  error: true
---

**NAMES:**


```{r}

```

Today you will be exploring different types of visualizations to uncover which species of penguins reside on different islands.

***This task is complex. It requires many different types of abilities. Everyone will be good at some of these abilities but nobody will be good at all of them. In order to solve this puzzle, you will need to use the skills of each member of your group.***

Some advice: If you aren't sure how to make a specific plot or how to customize
a look over the class text for this week and make use of the **ggplot2**
cheatsheet. 

# Getting Set-up

<!-- The partner whose family name starts first alphabetically starts as the  -->
<!-- Computer (typing and listening to instructions from the Coder! -->

Name of the person starting as the Computer (typing): 

# Exploring the Penguins Data

In this section you will explore the `penguins` dataset by creating multiple
types of plots. As you are creating you plots, please don't forget to (1) give
your plots reader friendly axes labels, and. (2) follow good code formatting!

## Step 0: Setup Chunk
In the empty code chunk at the top of the document, add the correct packages and label the code chunk as `setup`.  

- `tidyverse` or `ggplot2`  
- `palmerpenguins` 

## Step 1: Barchart

Use `ggplot2` to plot the frequency of each penguin `species`. 

```{r}
#| label: barchart

```


<!-- Change roles (Computer becomes Coder)! -->

## Step 2: Histogram or Density Curve

Use `ggplot2` to plot the distribution of bill lengths for the penguins included in the dataset.

```{r}
#| label: histogram-or-density

```

<!-- Change roles (Computer becomes Coder)! -->

## Step 3: Scatterplot

Use `ggplot2` to plot the relationship between the length of a penguin's bill (`bill_length_mm`) and the depth of their bill (`bill_depth_mm`).

```{r}
#| label: scatterplot

```

## Step 4: Add a Trend Line

Add a **linear** trend line to the scatterplot you made above! 


<!-- Change roles (Computer becomes Coder)! -->

## Step 5: Adding A Categorical Variable

Building off of the plot you made in Step 4, add an aesthetic to differentiate
the species of the penguins in the scatterplot by color.

```{r}
#| label: scatterplot-with-one-categorical-variable

```

<!-- Change roles (Computer becomes Coder)! -->

## Step 6: Adding Another Categorical Variable

Building off of the plot you made in Step 5, add the location of the penguins
( `island`) to your visualization. There may be more than one method to address
this, however, one method will more easily allow you to address the questions
below.

```{r}
#| label: scatterplot-with-two-categorical-variables

```


<!-- Change roles (Computer becomes Coder)! -->

# Canvas Quiz

Working together as a team, use the plots you both created to address the
following questions below, then submit your answers individually on Canvas:

1.  Which species of penguins is represented the **least** in the `penguins` data set?

>Insert Answer

2.  What is the most frequent bill length seen in these data? (Approximate to the nearest 5mm)  

>Insert Answer

3.  Which species of penguins are found on Biscoe Island?  

>Insert Answer

4.  Which species of penguins are found on Dream Island?  

>Insert Answer

5.  Which species of penguins are found on Torgersen Island?  

>Insert Answer



# Submission

Render the Quarto document to HTML. Submit the zipped file and the html separately for the quiz (so you will need to export this twice).  
