---
title: "Iteration"
format: html
embed-resources: true
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(repurrrsive)
library(broom)
```


```{r}
#| label: data-prep
#| message: false
#| warning: false

names(sw_people) <- sw_people  |>  map_chr("name")

# Toy data
set.seed(1000)
exams <- list(
  student1 = round(runif(10, 50, 100)),
  student2 = round(runif(10, 50, 100)),
  student3 = round(runif(10, 50, 100)),
  student4 = round(runif(10, 50, 100)),
  student5 = round(runif(10, 50, 100))
)

extra_credit <- list(0, 0, 10, 10, 15)
```



# Mapping

## purrr


### Your Turn 1

Run the code in the chunk. What does it do?

```{r}
#| label: calculate-mean
#View(exams)
exams |> map(mean)
```

### Example

```{r}
#| label: add-names
x <- c("hw1","hw2","hw3","hw4","hw5","hw6","hw7","hw8","hw9","hw10") 
exams |> map(set_names, nm = x)
```



### Example- map_dbl()

```{r}
#| label: mean-table
exams |> 
  map(mean)
```

### Example - map_dfr()

```{r}
#| label: data-frame-hw
exams |>
  map(set_names, nm = x) |>
  map(bind_rows)
```



## map shorthand


## Your Turn 2

Complete the code to apply the test below to every element of the list. Return the results as a _logical_ vector.

`length(<input>) == 10`

```{r}
#| label: ten-items
exams |>  ______(_________________)
```

## Your Turn 3

Recompute the grades by dropping the lowest score and then taking the mean of the rest, i.e. Return the results as a _double_ vector.

`(sum(<input>) - min(<input>)) / 9`

```{r}
#| label: drop-lowest
exams |> 
 
```


## Your Turn 4

Compute a final grade for each student, where the final grade is the average test score plus any `extra_credit` assigned to the student. Return the results as a double (i.e. numeric) vector.

`mean(<exams>) + <extra_credit>`

```{r}
#| label: extra-credit
exams  |>  
  _________(____________, ~________________)
```


# Putting it all together

Here is our data cleaning:

```{r}
#| label: data-cleaning-tb

#who is already a data set built into R packages

names(who) <- str_replace(names(who), "newrel", "new_rel")

tb <-  
  who |> 
  pivot_longer(cols = 5:60, names_to = "codes", values_to = "n") |>  #pivots data
  separate(codes, c("new", "type", "sexage"), sep = "_") |>  #separates labels into three variables
  select(-new, -iso2, -iso3) |> # removes unnecessary columns
  separate(sexage, c("sex", "age"), sep = 1) |>  # separates into sex and age column
  separate(age, c("age_l", "age_h"), sep = -2) |>  # separates age range into two columns
  unite(age_l:age_h, col = "age", sep = "-") |>   #puts age range back together with - between values
  mutate(age = ifelse(age == "-65", "65+", age)) |>  #fixes 65+ value
  drop_na(n) #drops missing values
```




Predict TB cases in 2020 for Afghanistan

```{r}
tb |> 
  filter(country == "Afghanistan") |>
  group_by(year) |> 
  summarise(n = sum(n)) |>
  ggplot(aes(x = year, y = n)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```


```{r}
tb |> 
  filter(country == "Afghanistan") |>
  group_by(year) |> 
  summarise(n = sum(n)) |> 
  with(lm(n ~ year)) |> 
  predict(newdata = tibble(year = 2020)) |> 
  round()
```

Predict TB cases in 2020 for each country

```{r}
tb_predictions <-
  tb |> 
  group_by(country, year) |> 
  summarise(n = sum(n), .groups = "drop_last") |> 
  group_by(country) |> 
  nest() |> 
  mutate(model = map(data, ~lm(n ~ year, data = .x))) |>
  mutate(pred = map_dbl(model, predict, tibble(year = 2020))) |>
  mutate(pred = round(pred))
```

Things that we can now do:

* Which countries will have the most cases?
* What does the data look like for India?
* How do the Korea's compare?


```{r}
tb_predictions |> 
  select(country, pred) |> 
  arrange(desc(pred))
```

```{r}
tb_predictions |>
  filter(country == "India") |> 
  unnest(data)
```

```{r}
tb_predictions |>
  filter(str_detect(country,"Korea")) |> 
  select(country, pred) 
```


***

Take Aways

* Use `View()` to inspect lists
* Use `pluck()`, `[[ ]]`, or `$` to extract values from lists
* Use purrr's `map()` functions to manipulate lists
* Lists can be a column in a data frame
* Tables are an organizational structure that you can manipulate.
