---
title: "Exploratory Data Analysis"
format: html
embed-resources: true
execute: 
  echo: true
---


```{r}
#| label: setup
library(tidyverse)
```


```{r}
#| label: load-babynames
babynames <- read_csv("babynames_500.csv")
```


# dplyr - `summarize()`

## Compute a table of summaries

```{r}
#| label: summarize-data

babynames |> 
  summarize(total = sum(n),
            max = max(n))
```


## Your Turn 1

Complete the code below to extract the rows where `name == "Lincoln"`. Then use `summarise()` and `sum()` and `min()` to find:

1. The total number of children named Lincoln
2. The first year Lincoln appeared in the data

*(Hint: Be sure to remove each `_` before running the code)*

```{r}
#| label: popularity-lincoln

babynames ___ 
  filter(_______________________) ___
  ___________(total = ________, 
              first = _______)
```



## The `n()` Function
The `n()` function does not need any input. It simply counts the number of rows in the data frame or tibble.  

```{r}
#| label: n-rows

babynames |> 
  summarize(row_n = n())
```

## The `n_distinct()` Function

```{r}
#| label: n-distinct-rows

babynames |> 
  summarize(row_n = n(),
            name_n = n_distinct(name))
```

## Name Popularity 

```{r}
#| label: name-popularity

babynames |> 
  filter(name == "Judith",
         sex == "F") |> 
  summarize(total = sum(n))
```
# dplyr - `group_by()`

**NOTE** The `group_by()` function is never used alone. It is always used before either `summarize()` or `mutate()` to communicate to those functions how to perform thier calculations.  

```{r}
#| label: create-pollution
#| echo: false

pollution <- tribble(
  ~city,   ~size, ~amount,
  "New York", "large", 23,
  "New York", "small", 14,
  "London", "large", 22,
  "London", "small", 16,
  "Beijing", "large", 121,
  "Beijing", "small", 56
)
```



```{r}
#| label: summarize-pollution

pollution |> 
  summarize(mean = mean(amount),
            sum = sum(amount),
            n = n())
```

```{r}
#| label: summarize-city-pollution

pollution |> 
  group_by(city) |> #only new line of code
  summarize(mean = mean(amount),
            sum = sum(amount),
            n = n())
```


```{r}
#| label: summarize-babynames

babynames |> 
  group_by(sex) |> 
  summarize(total = sum(n))

```





## Your Turn 2

Use `group_by()`, `summarise()`, and `arrange()` to display the ten most popular names. Compute popularity as the *total* number of children of a single gender given a name.

*(Hint: Be sure to remove each `_` before running the code)*

```{r}
#| label: top-10-names

babynames |> 
  _______(name, sex) |>  
  _______(total = _____(n)) |> 
  _______(desc(_____)) |> 
  head(10)
```


Go ahead and add one the `ggplot` code to make the visualization.


# dplyr - `mutate()`
Use `mutate()` to create NEW columns/variables within your data set.  


```{r}
#| label: percentage-baby

babynames |> 
  mutate(percent = round(prop*100, 2)) |> 
  relocate(percent, .before = proportion)
```
```{r}
#| label: percent-2-baby

babynames |> 
  mutate(percent = round(prop*100, 2),
         nper = round(percent)) |> 
  relocate(percent:nper, .before = proportion)
```

## Your Turn 3

Fill in the blanks to: 
Rank each name _within its year and sex_. 

*(Hint: Be sure to remove each `_` before running the code)*

```{r}
#| label: ranking-names

babynames |> 
  select(name, year, sex, prop) |> 
  group_by(____, ____) |> 
  ______(rank = min_rank(____(prop))) |> 
  arrange(year, sex, rank)
```


***

# Take aways

* Make tables of summaries with `summarise()`  
* Make new variables, with `mutate()`  
* Do groupwise operations with `group_by()`

* Connect operations with `%>%` or `|>`  

