---
title: "Join Data with dplyr"
format: html
embed-resources: true
---

```{r setup}
#| label: setup
library(tidyverse)
library(nycflights13)
```


```{r}
#| label: toy-data

band <- tribble(
   ~name,     ~band,
  "Mick",  "Stones",
  "John", "Beatles",
  "Paul", "Beatles"
)

instrument <- tribble(
    ~name,   ~plays,
   "John", "guitar",
   "Paul",   "bass",
  "Keith", "guitar"
)

instrument2 <- tribble(
    ~artist,   ~plays,
   "John", "guitar",
   "Paul",   "bass",
  "Keith", "guitar"
)
```

# nycflights13

<!-- Note - we don't want to evaluate the `View` functions when we render the document. -->

```{r}
#| eval: false
View(flights)
```

```{r}
#| eval: false
View(airlines)
```


# mutating joins

```{r}
#| label: left-join
#keeps the data in band and adds matching columns from instrument
band |> left_join(instrument, by = "name")
```


```{r}
#| label: right-join
#keeps the data in instrument and adds matching columns from band
band |> right_join(instrument, by = "name")
```


```{r}
#| label: full-join
#keeps columns from both band and instrument
band |> full_join(instrument, by = "name")
```


```{r}
#| label: inner-join
#keeps only that which matches both band and instrument
band |> inner_join(instrument, by = "name")
```


## Your Turn 1a

Which airlines had the largest arrival delays? Complete the code below.

1. Join `airlines` to `flights`

*(Hint: Be sure to remove each `_` before running the code)*

```{r}
#| label: match-airlines-fligths

flights |>
  drop_na(arr_delay) |>
  ___________(airlines, by = "_________")  |>
  head() 
```

## Your Turn 1b

Which airlines had the largest arrival delays? Complete the code below.

1. Join `airlines` to `flights`
2. Compute and order the average arrival delays by airline. Display full names, no codes.

*(Hint: Be sure to remove each `_` before running the code)*

```{r}
#| label: avg-arrival-delay

flights |>
  drop_na(arr_delay) |>
  __________________  |>
  group_by(_________) |>
  __________________  |>
  arrange(__________) 
```

## Your Turn 2a

Join `flights` and `airports` by `dest` and `faa`. 


```{r}
#| label: join-flights-airports

flights |> 
  drop_na(arr_delay) |>
  ______(airports, __________________) |>
  head()
```


## Your Turn 2b

Join `flights` and `airports` by `dest` and `faa`. 

Then for each `name`, compute the `distance` from NYC and the average `arr_delay`.  *Hint: use `first()` to get the first value of distance.*

Order by average delay, worst to best.

*(Hint: Be sure to remove each `_` before running the code)*

```{r}
#| label: avg-distance

flights |> 
  drop_na(arr_delay) |>
  ______(airports, __________________) |>
  group_by(name) |>
  _________(distance = _____________, delay = _____________) |>
  ______(_____(delay))

```



# filtering joins

```{r}
#| label: semi-join
band |> semi_join(instrument, by = "name")
```


```{r}
#| label: anti-join
band |> anti_join(instrument, by = "name")
```

## Your Turn 3

How many airports in `airports` are serviced by flights in `flights`? (i.e. how many places can you fly to direct from New York?) 

Notice that the column to join on is named `faa` in the **airports** data set and `dest` in the **flights** data set.

```{r}
#| label: airports-nyc
__________ |>
  _____________________________ |>
  select(faa)
```


# Other Functions  

```{r}
#| label: distinct

instruments |> 
  distinct(plays)
```


# Take aways

* `left_join()` retains all cases in _left_ data set
* `right_join()` retains all cases in _right_ data set
* `full_join()` retains all cases in _either_ data set
* `inner_join()` retains _only_ cases in _both_ data sets

* `semi_join()` extracts cases that _have_ a match
* `anti_join()` extracts cases that _do not have_ a match 


