---
title: "Tidy Data"
format: html
---

```{r}
#| label: setup
library(tidyverse)
```


```{r}
#| label: create-toy-data

# Toy data
cases <- tribble(
  ~Country, ~"2011", ~"2012", ~"2013",
  "FR",    7000,    6900,    7000,
  "DE",    5800,    6000,    6200,
  "US",   15000,   14000,   13000
)

pollution <- tribble(
  ~city, ~size, ~amount,
  "New York", "large",      23,
  "New York", "small",      14,
  "London", "large",      22,
  "London", "small",      16,
  "Beijing", "large",     121,
  "Beijing", "small",     121
)

names(who) <- stringr::str_replace(names(who), 
                                   "newrel", 
                                   "new_rel")
```

Run the code below to look at the tables from the presentation 

```{r}
table1
table2
table3
table4a
table4b
table5
```

# tidyr


## Your Turn 1

On a sheet of paper, draw how the cases data set would look if it had the same values grouped into three columns: **country**, **year**, **n**

```{r}
cases |> 
  pivot_longer(names_to = "year", 
               values_to = "n", 
               cols = 2:4)
```



## Your Turn 2
Try the following codes - what do they do? (annotate them with a comment)

```{r}
cases |> 
  pivot_longer(names_to = "year",
               values_to = "n", 
               cols = 2:4) #comment

cases |> 
  pivot_longer(names_to = "year",
               values_to = "n", 
               cols = c("2011", "2012", "2013")) #comment

cases |> 
  pivot_longer(names_to = "year",
               values_to = "n", 
               cols = starts_with("201")) #comment

cases |> 
  pivot_longer(names_to = "year",
               values_to = "n", 
               cols = -Country) #comment
```



## Your Turn 3

Use `pivot_longer()` to reorganize `table4a` into three columns: **country**, **year**, and **cases**.

```{r}


```


## Your Turn 4

On a sheet of paper, draw how this data set would look if it had the same values grouped into three columns: **city**, **large**, **small**




## Your Turn 5

Use `pivot_wider()` to reorganize `table2` into four columns: **country**, **year**, **cases**, and **population**.

```{r}

```

# who


## Your Turn 6

Gather the 5th through 60th columns of `who` into columns named `codes` and `n`.

Then select just the `country`, `year`, `codes` and `n` variables.

```{r}
who |>
  pivot_longer(names_to = _____, 
               values_to = ____, 
               cols = 5:60) |>
  select(-iso2, -iso3)

```



Run the following code
```{r}
who |>
  pivot_longer(names_to = "codes", 
               values_to = "n", 
               cols = 5:60) |>
  select(-iso2, -iso3) |>
  separate(col = codes, 
           into = c("new", "type", "sexage"), 
           sep = c(4, 7))
```




## Your Turn 7

Separate the `sexage` column into `sex` and `age` columns.

*(Hint: Be sure to remove each `_________` before running the code)*

```{r}
who |>
  pivot_longer(names_to = "codes", 
               values_to = "n", 
               cols = 5:60) |>
  select(-iso2, -iso3) |>
 separate(col = codes, 
          into = c("new", "type", "sexage"), 
          sep = "_") |>
  select(-new) |> 
  _______________________________
```


***

# Take Aways

Data comes in many formats but R prefers just one: _tidy data_.

A data set is tidy if and only if:

1. Every variable is in its own column
2. Every observation is in its own row
3. Every value is in its own cell (which follows from the above)

What is a variable and an observation may depend on your immediate goal.

