---
title: "Functions"
format: html
embed-resources: true
execute: 
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
```


```{r}
#| label: fake-data
df <- tibble::tibble(
  a = c(rnorm(9), -99),
  b = c(-999, -99, rnorm(8)),
  c = c(0, rnorm(9)),
  d = rnorm(10)
)
```


Rescaling is a process to make all data have the same overall scale while maintaining characteristics of the spread within the data.

```{r}
((df$a - min(df$a)) / (max(df$a) - min(df$a)))
((df$b - min(df$b)) / (max(df$b) - min(df$b)))
((df$c - min(df$c)) / (max(df$c) - min(df$c)))
((df$d - min(df$d)) / (max(df$d) - min(df$c)))
```


Write a function to automate the calculation of the rescale:

```{r}
rescale01 <- function(x) {
  

}

rescale01(df$a)
rescale01(df$b)
rescale01(df$c)
rescale01(df$d)
```


## Your Turn 1

Rerun the tibble to revert back to the original data.  Write a function to replace -99 with NA, called `fix_missing()`

```{r}
df <- tibble::tibble(
  a = c(rnorm(9), -99),
  b = c(-999, -99, rnorm(8)),
  c = c(0, rnorm(9)),
  d = rnorm(10)
)


#Here are the repeated calculations
df$a[df$a == -99] <- NA
df$b[df$b == -99] <- NA
df$c[df$c == -99] <- NA
df$d[df$d == -99] <- NA

```


```{r}
fix_missing <- function() {
  
  
  
}
```



Now let's write a function that allows us to easily adapt our rescaling to include a range (min, max) to modify the data.
```{r}
#Rescale to [0, 1]
0 + (1 - 0) * ((df$a - min(df$a)) / (max(df$a) - min(df$a)))

# Rescale to [-1, 1]
-1 + (1 - -1) * ((df$b - min(df$b)) / (max(df$b) - min(df$b)))

# Rescale to [0, 10]
0 + (10 - 0) * ((df$c - min(df$c)) / (max(df$c) - min(df$c)))

```


```{r}
rescale <- function(x, min =  , max = ) {
  

}

rescale(df$a)
rescale(df$b, min = -1, max = 1)
rescale(df$c, max = 10)

```




## Your Turn 2
Expand your function from 'Your Turn 1' to allow for any possible missing value type.
```{r}



```



Now, let's make our functions pipeable 

```{r}
rescale <- function(data, x, min = 0, max = 1) {
  data |>
    summarize(rescale = min +
      (max - min) * (({{ x }} - min({{ x }},
        na.rm = TRUE
      )) / (max({{ x }},
        na.rm = TRUE
      ) - min({{ x }}, na.rm = TRUE))))
}

rescale(data = df, x = a)

```




# Take Aways

To write a function, 

1. Write code that solves the problem for a real object  
2. Wrap the code in `function(){}` to save it  
3. Add the name of the real object as the function argument 

This sequence will help prevent bugs in your code (and reduce the time you spend correcting bugs). 
