---
title: "PA 3: Identify the Mystery College 🏫"
subtitle: "Data Wrangling with `dplyr`"
author: "Names of Your Group!"
format: html
embed-resources: true
code-tools: true
toc: true
editor: source
execute: 
  error: true
  echo: true
  eval: false
  message: false
  warning: false
---

Today you will use the `dplyr` package to clean some data. We will then use that
cleaned data to figure out what college Ephelia has been accepted to.

***This task is complex. It requires many different types of abilities. Everyone will be good at some of these abilities but nobody will be good at all of them. In order to produce the best product possible, you will need to use the skills of each member of your group.***

# Part One: Data Import and Cleaning

This section will clean today's data so that you can use it more easily in Part
Two.

<!-- The partner whose birthday is closest to January 1st starts as the  -->
<!-- Developer (typing and listening to instructions from the Coder! -->

## Data Download & Package Loading

First, we declare our package dependencies and load the data.

*The data loading function `read_csv()` will give you an outpouring of helpful information about the dataset. If you do not see the word "error", there is nothing to be concerned about.*

```{r}
#| warning: false
#| message: false
#| label: load-data-packages

library(tidyverse)

colleges <- read_csv("https://www.dropbox.com/s/bt5hvctdevhbq6j/colleges.csv?dl=1")
```

Take a look at the variables in your downloaded data by running the following
code. This tells reports the data type for each column in the dataset.

```{r}
glimpse(colleges)
```

## Data Cleaning

Now we will clean the data. Alas, each of the R chunks in this section will
cause an error and / or do the desired task incorrectly. Even the chunks that
run without error **are not** correct! You will need to find the mistake and
correct it to complete the intended action.

**Step 1:** There are too many variables in this data set. We don't need all of
them. Narrow your data set down to only:

-   `INSTNM` name of the institution
-   `CITY` city, `STABBR` state, and `ZIP` ZIP code of the institution
-   `ADM_RATE` admissions rate
-   `SAT_AVG` average SAT score
-   `UGDS` number of undergraduate students
-   `TUITIONFEE_IN` in- and `TUITIONFEE_OUT` out-of-state tuition
-   `CONTROL` Whether the school is public or private
-   `REGION` region of the school.

```{r}
#| label:  select-vars

colleges_clean1 <- colleges | >  
  select(INSTNM, 
         CITY, 
         STABBR, 
         ZIP,
         ADM_RATE, 
         SAT_AVG, 
         UGDS,
         TUITIONFEE_IN, 
         TUITIONFEE_OUT
         CONTROL,
         REGION) 
```

**Step 2:** Remove the schools that are private and for-profit (category 3).

```{r}
#| label: filter-private-profit

colleges_clean2 <- colleges_clean1 |> 
  filter(CONTROL == 3)
```

**Step 3:** Adjust the appropriate variables to be numeric, using `as.numeric()`.

```{r}
#| label: mutate-numeric

colleges_clean3 <- colleges_clean2 |> 
  mutate(TUITIONFEE_IN  = numeric(TUITIONFEE_IN),
         TUITIONFEE_OUT = numeric(TUITIONFEE_OUT),
         SAT_AVG        = numeric(SAT_AVG),
         UGDS           = numeric(UGDS),
         ADM_RATE       = numeric(ADM_RATE)) 
```

<!-- Swap roles -- Developer becomes Coder, Coder becomes Developer! -->

**Step 4:** Adjust the appropriate variables to be factors, using `as.factor()`.

```{r}
#| label: mutate-factor

colleges_clean4 <- colleges_clean3 |>
  mutate(CONTROL = as.character(CONTROL),
         REGION  = as.character(REGION))
```

**Step 5:** Create a new variable called `TUITION_DIFF` which contains the
difference between out-of-state and in-state costs.

```{r}
#| label: add-differences-between-in-state-out-of-state

colleges_clean5 <- colleges_clean4 |> 
    TUITION_DIFF = TUITIONFEE_OUT - TUITIONFEE_IN
```

**Step 6:** Remove every row with missing data.

```{r}
#| label: drop-na

colleges_clean6 <- colleges_clean5 |> 
  drop.na()
```

At this point we've made six different datasets, but notice each of these steps
started with `colleges_clean <- colleges_clean |>`. That is pretty redundant! 
Instead, we could have performed all these tasks as one long "pipeline."

**Step 7:** Combine your (fixed) code chunks into a **single** code chunk that
carries out all of the steps necessary to clean the data. 

*Think about coding efficiency -- you __should not__ have multiple calls to the __same__ function!*

```{r}
#| label: all-steps-one-pipeline
# Code combining ALL of your previous steps into ONE pipeline

colleges_clean <- colleges |>
  
```

<!-- Swap roles -- Developer becomes Coder, Coder becomes Developer! -->

# Part Two: Identify the Mystery College

Wow! Your best friend Ephelia has been accepted to her top choice college!
Unfortunately, Ephelia is a very mysterious person, and she won't tell you
directly which college this is. You'll have to use her clues to figure out which
school is her dream school.

**Clues:**

1.  This college is located in Region 7.

2.  This college's admission rate is above the median rate for the region.

3.  This college **does not** charge the same for in- and out-of-state tuition.

4.  The average SAT score of this college is an odd number.

<!-- Swap roles -- Developer becomes Coder, Coder becomes Developer! -->

5.  This college **is not** in Idaho.

6.  Less than 10,000 undergraduates attend this college.

7.  The college **is not** in the state where Dr. Theobold received a Ph.D.

8.  Of the five options remaining at this step, Ephelia will attend the college
that has the lowest admission rate.

```{r}
#| label: identify-college 


```

Submit the college Ephelia will attend to the Canvas Quiz.
