---
title: "PA 3 Task Card -- Finding Ephelia's College"
editor: source
format: docx
---

Today you will use the `dplyr` package to clean some data. We will then use that
cleaned data to figure out what college Ephelia has been accepted to.

***This task is complex. It requires many different types of abilities. Everyone
will be good at some of these abilities but nobody will be good at all of them.
In order to solve this puzzle, you will need to use the skills of each member of
your group.***

## Group Roles

When you are the **Developer**, you will type into the Quarto document in
RStudio. However, you **do not** type your own ideas. Instead, you type what 
the Coder tells you to type. You are permitted to ask the Coder clarifying
questions, and, if both of you have a question, you are permitted to ask the
professor. You are expected to run the code provided by the Coder and, if 
necessary, to work with the Coder to debug the code. Once the code runs, you
are expected to collaborate with the Coder to write code comments that describe
the actions taken by your code.

When you are the **Coder**, you are responsible for reading the 
instructions / prompts and directing the Developer what to type in the Quarto
document. You are responsible for managing the resources your group has
available to you (e.g., cheatsheet, textbook). If necessary, you should work
with the Developer to debug the code you specified. Once the code runs, you
are expected to collaborate with the Developer to write code comments that
describe the actions taken by your code.

## Group Norms

Remember, your group is expected to adhere to the following norms:

1.  Think and work together. Do not divide the work.
2.  You are smarter together.
3.  Be open minded. 
4.  No cross-talk with other groups.
5.  Communicate with each other! 

{{< pagebreak >}}

## Piping

You can choose whether to: 

declare your data as the first argument of the function 
```{r}
#| echo: true
#| eval: false
#| code-line-numbers: false
#| label: not-piped-data

filter(.data = colleges, REGION == 5) |>
  mutate(TUITION_DIFF = TUITIONFEE_OUT - TUITIONFEE_IN)
```

***or***

pipe your data into the first argument of the function
```{r}
#| echo: true
#| eval: false
#| code-line-numbers: false
#| label: pipe-data

colleges |>
  filter(REGION == 5) |>
  mutate(TUITION_DIFF = TUITIONFEE_OUT - TUITIONFEE_IN)
```