---
title: "Lab 2: Exploring Rodents with `ggplot2`"
author: "Instructions"
format: html
editor: source
number-sections: true
number-depth: 2
toc: true
code-tools: true
execute:
  echo: true
  eval: false
  message: false
  warning: false
  error: true
resources: 
  - ../student/lab-2-student.qmd
  - data/surveys.csv
---

# Part One: Set-up

## Accessing the Lab & Data

Download the template Lab 2 Quarto file here: <a href="../student/lab-2-student.qmd" download>`lab-2-student.qmd`</a> 

Download the data for the lab here: [`surveys.csv`](data/surveys.csv)


::: {.callout-important}
Be sure to save the file in the Lab 2 folder, inside your Week 2 folder, 
inside your STAT 331 folder! 
:::

## Using External Resources

Part of learning to program is learning from a variety of resources. Thus, I
expect you will use resources that you find on the internet. There is, however,
an important balance between copying someone else's code and
*using their code to learn*. Therefore, if you use external resources, I want to
know about it. 

- If you used Google, you are expected to "inform" me of any resources you used
by **pasting the link to the resource in a code comment next to where you used
that resource**.

- If you used ChatGPT, you are expected to "inform" me of the assistance you
received by (1) indicating somewhere in the problem that you used ChatGPT 
(e.g., below the question prompt or as a code comment), and (2) downloading and
attaching the `.txt` file containing your **entire** conversation with ChatGPT.

Additionally, you are permitted and encouraged to work with your peers as you
complete lab assignments, but **you are expected to do your own work**. Copying
from each other is cheating, and letting people copy from you is also cheating.
Please don't do either of those things.

# Part Two: Data Context

The questions in this lab are noted with numbers and boldface. Each question
will require you to produce code, whether it is one line or multiple lines.

This document is quite plain, meaning it does not have any special formatting.
As part of your demonstration of creating professional looking Quarto documents,
I would encourage you to spice your documents up (e.g., declaring execution
options, specifying how your figures should be output, formatting your code
output, etc.).

## Setup

In the code chunk below, load in the packages necessary for your analysis. You
should only need the `tidyverse` package for this analysis.

```{r}
#| label: setup

```

## Data Context

The Portal Project is a long-term ecological study being conducted near Portal, 
AZ. Since 1977, the site has been used to study the interactions among rodents,
ants, and plants, as well as their respective responses to climate. To study the
interactions among organisms, researchers experimentally manipulated access to
24 study plots. This study has produced over 100 scientific papers and is one of
the longest running ecological studies in the U.S.

We will be investigating the animal species diversity and weights found within
plots at the Portal study site. The data are stored as a comma separated value
(CSV) file. Each row holds information for a single animal, and the columns
represent:

| Column            | Description                        |
|-------------------|------------------------------------|
| `record_id`       | Unique ID for the observation      |
| `month`           | month of observation               |
| `day`             | day of observation                 |
| `year`            | year of observation                |
| `plot_id`         | ID of a particular plot            |
| `species_id`      | 2-letter code                      |
| `sex`             | sex of animal ("M", "F")           |
| `hindfoot_length` | length of the hindfoot in mm       |
| `weight`          | weight of the animal in grams      |
| `genus`           | genus of animal                    |
| `species`         | species of animal                  |
| `taxon`           | e.g. Rodent, Reptile, Bird, Rabbit |
| `plot_type`       | type of plot                       |

## Reading the Data into `R`

**1. Using the `read_csv()` function, write the code necessary to load in the `surveys.csv` dataset (stored in the data folder). For simplicity, name the data `surveys`.**

```{r}
#| label: load-data

```

**2. What are the dimensions (# of rows and columns) of these data?**

<!-- This question requires you to write code! You will need to make a new code chunk! -->


**3. What are the data types of the variables in this dataset?**

<!-- This question requires you to write code! You will need to make a new code chunk! -->


# Exploratory Data Analysis with **`ggplot2`**

`ggplot()` graphics are built step by step by adding new elements. Adding layers
in this fashion allows for extensive flexibility and customization of plots.

To build a `ggplot()`, we will use the following basic template that can be used
for different types of plots:

```{r}
#| eval: false
#| label: ggplot-template-code

ggplot(data = <DATA>,
       mapping = aes(<VARIABLE MAPPINGS>)) +
  <GEOM_FUNCTION>()
```

Let's get started!

## Scatterplot

**4. First, let's create a scatterplot of the relationship between `weight` (on
the $x$-axis) and `hindfoot_length` (on the $y$-axis).**

```{r}
#| label: scatterplot

```

We can see there are **a lot** of points plotted on top of each other. Let's try and modify this plot to extract more information from it.

**5. Let's add transparency (`alpha`) to the points, to make the points more transparent and (possibly) easier to see.**

<!-- Do not write code here! Add code to the scatterplot you created for question 4! -->

Despite our best efforts there is still a substantial amount of overplotting occurring in our scatterplot. Let's try splitting the dataset into smaller
subsets and see if that allows for us to see the trends a bit better.

**6. Facet your scatterplot by `species`.**

<!-- Do not write code here! Continue to modify the scatterplot you created for question 4! -->

**7. No plot is complete without axis labels and a title. Include reader friendly labels and a title to your plot.**

<!-- Do not write code here! Continue to modify the scatterplot you created for question 4! -->

It takes a larger cognitive load to read text that is rotated. It is common practice in many journals and media outlets to move the $y$-axis label to the top of the graph under the title.

**8. Specify your $y$-axis label to be empty and move the $y$-axis label into the subtitle.** 

<!-- Do not write code here! Continue to modify the scatterplot you created for question 4! -->

## Boxplots

```{r}
#| label: boxplot


```

**10. Create side-by-side boxplots to visualize the distribution of `weight` (y-axis) within each `species` (x-axis).**

A fundamental complaint of boxplots is that they do not plot the raw data. However, with **ggplot** we can add the raw data points on top of the boxplots!

**11. Add another layer to your previous plot that plots each observation.**

<!-- Do not write code here! Add code to the boxplot you created for question 10! -->

Alright, this should look less than optimal. Your points should appear rather stacked on top of each other. To make them less stacked, we need to jitter them
a bit, using `geom_jitter()`.

**12. Remove the previous layer and include a `geom_jitter()` layer instead.**

<!-- Do not write code here! Continue to modify the plot you made for question 10! -->

That should look a bit better! But its really hard to see the points when everything is black.

**13. Set the `color` aesthetic in `geom_jitter()` to change the color of the points and add set the `alpha` aesthetic to add transparency.**

<!-- Do not write code here! Continue to modify the plot you made for question 10! -->

::: callout-note
You are welcome to use whatever color you wish! Some of my favorites are
`"orange3"` and `"steelblue"`.
:::

Great! Now that you can see the points, you should notice something odd: there
are two colors of points still being plotted. Some of the observations are being plotted twice, once from `geom_boxplot()` as outliers and again from `geom_jitter()`!

**14. Inspect the help file for `geom_boxplot()` and see how you can remove the outliers from being plotted by `geom_boxplot()`. Make this change in your code!**

<!-- Continue to modify the plot you made for question 10! -->

Some small changes can make **big** differences to plots. One of these changes
are better labels for a plot's axes and legend.

**15. Modify the $x$-axis and $y$-axis labels to describe what is being plotted. Be sure to include any necessary units! You might also be getting overlap in the species names -- use `theme(axis.text.x = ____)` or `theme(axis.text.y = ____)` to turn the species axis labels 45 degrees.**

<!-- Continue to modify the plot you made for question 10! -->

Some people (and journals) prefer for boxplots to be stacked with a specific orientation! Let's practice changing the orientation of our boxplots.

**16. Now copy-paste your boxplot code you've been adding to above. Flip the orientation of your boxplots. The variable that was mapped to the x-axis should now be included on the y-axis and visa versa.**

```{r}
#| label: rotated-boxplot

```

Notice how vertically stacked boxplots make the species labels more readable
than horizontally stacked boxplots (even when the axis labels are rotated). 
This is good practice!

# Conecting Visualizations with Statistical Analyses

Exploratory Data Analysis (EDA) is always a great start to investigating a 
dataset. Can we see a relationship between rodent weight and hindfoot length?
How does rodent weight differ between species? After performing EDA, we can then
conduct appropriate statistical analyses to formally investigate what we have
seen.

In this section, we are going to conduct a one-way analysis of variance (ANOVA)
to compare mean weight between the fourteen different species of rodents.

::: callout-tip
## Refresher on one-way ANOVA

While a second course in statistics is a pre-requisite for this class, you may
want a [refresher](https://openintro-ims.netlify.app/inference-many-means) on
conducting a one-way ANOVA.
:::

I have outlined the null and alternative hypotheses we will be testing:

::: columns
::: {.column width="45%"}
::: {.callout-note icon="false"}
## Null

The population mean weight is the same between **all** fourteen rodent species.
:::

:::

::: {.column width="5%"}
:::

::: {.column width="45%"}
::: {.callout-tip icon="false"}
## Alternative

**At least one** rodent species has a different population mean weight.
:::
:::
:::

Look up the help documentation for `aov()`.

**17. Using `aov()`, complete the code below to carry out the analysis.** You 
might want to look up the help documentation for `aov()` to know what inputs this 
function expects.

```{r}
#| label: anova

species_mod <- aov()

summary(species_mod)
```

**18. Based on the results of the ANOVA F-test, draw a conclusion in context of the hypotheses. Make sure to cite appropriate output from above.**


# Lab 2 Submission

For Lab 2 you will submit **only** your HTML file. Your HTML file is required to have the following specifications in the YAML options (at the top of your document):

-   have the plots embedded (`embed-resources: true`)
-   include your source code (`code-tools: true`)
-   include all your code and output (`echo: true`)

**If any of the options are not included, your Lab 2 or Challenge 2 assignment will receive an "Incomplete" and you will be required to submit a revision.**

In addition, your document should not have any warnings or messages output in your HTML document. **If your HTML contains warnings or messages, you will receive an "Incomplete" for document formatting and you will be required to submit a revision.**
